\documentclass[10pt,letterpaper]{article}

\usepackage[margin=0.55in]{geometry}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage[hidelinks]{hyperref}
\usepackage{fontawesome}
\usepackage{tabularx}
\usepackage{array}

\hypersetup{pageanchor=false}
\pagenumbering{gobble}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0pt}

\titleformat{\section}
  {\large\bfseries\uppercase}
  {}
  {0pt}
  {}
  [\titlerule]

\titlespacing{\section}{0pt}{6pt}{4pt}

\setlist[itemize]{
    leftmargin=*,
    itemsep=1pt,
    topsep=2pt,
    parsep=0pt,
    partopsep=0pt
}

\newcommand{\resumeSubheading}[4]{
  \textbf{#1} \hfill #2 \\
  \textit{#3} \hfill \textit{#4}
}

\newcommand{\resumeProjectHeading}[2]{
  \textbf{#1} \hfill \textit{#2}
}

\begin{document}

% ===================== HEADER =====================

\begin{center}
    {\LARGE \textbf{RICARDO ALEXANDER BARTRA QUISPE}} \\[3pt]
    Lima, Peru \;|\; \href{mailto:ricardo.bartra@pucp.edu.pe}{ricardo.bartra@pucp.edu.pe} \;|\; +51 919 503 758 \\
    \href{https://www.linkedin.com/in/ribartra}{linkedin.com/in/ribartra} \;|\;
    \href{https://github.com/ribartra}{github.com/ribartra}
\end{center}

% ===================== SUMMARY =====================

\section{Professional Summary}

AI and backend engineer focused on document intelligence, OCR validation, microservice architectures, and applied machine learning. Built production-oriented prototypes for Peruvian identity and education documents, human-in-the-loop OCR review, SAT practice workflows, legal/regulatory scraping, and image-recognition APIs. Strong practical stack across Python/FastAPI, Rust/Axum, Java/Spring Boot, TypeScript/Next.js, PostgreSQL, Redis, Docker Compose, Prisma, SQLAlchemy, OCR engines, PDF parsing, computer vision, and metrics-driven evaluation.

% ===================== SKILLS =====================

\section{Skills}

\textbf{Languages:} Python, Rust, Java, TypeScript, JavaScript, SQL, Bash \\

\textbf{Backend:} FastAPI, Axum, Express, Spring Boot, REST APIs, async services, JWT, webhooks, service orchestration, API gateways \\

\textbf{AI, OCR \& Data:} OCR, document understanding, computer vision, OpenCV, Docling, RapidOCR, GLM-OCR/Ollama, YOLO, pdfplumber, PyMuPDF, pandas, NumPy, SciPy, embeddings, RAG concepts \\

\textbf{Infrastructure:} Docker, Docker Compose, PostgreSQL, Redis Streams, SQLite, MySQL, S3 storage, Prometheus, GitHub Actions \\

\textbf{Frontend:} Next.js App Router, React, Tailwind CSS, Radix UI, TanStack Table, React Hook Form, Zod, NextAuth, Prisma \\

\textbf{Evaluation \& Optimization:} Benchmarking, extraction metrics, quality gates, fuzzy matching, regression analysis, model/service latency, ACO, PSO, ABC, combinatorial optimization

% ===================== EXPERIENCE =====================

\section{Experience}

\resumeSubheading
{AI / Backend Engineering Contributor}
{[Start Date] -- Present}
{Independent, Academic and Project-Based Work}
{Lima, Peru}

\begin{itemize}
    \item Built OCR and document-validation services for DNI, COE, CLA, CNR, FRP and related document families using Python, FastAPI, OCR engines, PDF parsers, image preprocessing, fuzzy matching, and structured JSON outputs.
    \item Designed a Docker Compose backend for document processing with Rust APIs, Python workers, PostgreSQL, Redis queues, review services, rule configuration, status tracking, authentication, rate limiting, and observability endpoints.
    \item Developed Rust services with Axum, Tokio, Redis, JWT, Prometheus, tracing, and SQL-backed service boundaries for gateway, ingestion, queue dispatching, rules, and status APIs.
    \item Implemented human-in-the-loop review workflows with FastAPI, SQLAlchemy async, Alembic, Pydantic, PostgreSQL, S3-compatible storage, and manual QA states for OCR pipelines.
    \item Created OCR benchmarks and validation reports measuring extraction quality, image resolution impact, execution time, and field-level match quality across document formats.
    \item Built web application workflows with Next.js, React, TypeScript, Prisma, NextAuth, Tailwind, Radix UI, Zod, and TanStack Table for role-based portals, assignments, testing flows, and administrative dashboards.
\end{itemize}

\resumeSubheading
{Research Assistant -- Applied AI and Optimization}
{[Start Date] -- Present}
{[University / Research Lab / Project Name]}
{Lima, Peru}

\begin{itemize}
    \item Designed experiments for applied AI, document understanding, retrieval-augmented generation, model evaluation, quantization, and lightweight deployment.
    \item Implemented and compared metaheuristic approaches including Ant Colony Optimization with Time Windows, Artificial Bee Colony, and Particle Swarm Optimization for routing, scheduling, and constrained search problems.
    \item Built reproducible evaluation loops with train/test separation, metric computation, confusion matrices, regression metrics, custom objective functions, and comparative experiment summaries.
    \item Produced technical documentation, architecture notes, implementation guides, API references, and benchmark reports for AI and backend systems.
\end{itemize}

% ===================== PROJECTS =====================

\section{Selected Projects}

\resumeProjectHeading
{OCR Document Processing Platform}{Rust, Python, FastAPI, Axum, PostgreSQL, Redis, Docker}

\begin{itemize}
    \item Designed a multi-service OCR backend with API gateway, ingest API, status API, rules API, queue service, review API, and specialized detect/extract workers.
    \item Implemented document-family workflows for DNI, CNR, COE, FRP, CLA and LN using OCR, layout detection, PDF parsing, validation rules, and worker-based processing.
    \item Used Redis Streams for dispatching, PostgreSQL for service-owned persistence, Docker Compose for orchestration, Prometheus for metrics, and JWT/internal tokens for service access.
    \item Added human review integration for low-confidence or manually validated OCR results.
\end{itemize}

\resumeProjectHeading
{DNI and RENIEC Document Validators}{Python, FastAPI, OpenCV, Docling, GLM-OCR, Ollama}

\begin{itemize}
    \item Built a DNI validation pipeline with image quality checks, resizing, OCR by text boxes, dual-format field extraction, exact DNI matching, fuzzy name matching, and JSON/HTML reporting.
    \item Supported multiple OCR paths including GLM-OCR through Ollama and Docling, with benchmarking for resolution, latency, and extraction quality.
    \item Extended document extraction patterns to RENIEC and related Peruvian document validation workflows using structured outputs and test datasets.
\end{itemize}

\resumeProjectHeading
{COE Certificate Extraction and Validation}{Python, FastAPI, pdfplumber, pandas, NumPy}

\begin{itemize}
    \item Built extraction logic for Peruvian Certificates of Studies across DCN 2009, CNEB 2016, and transition formats.
    \item Combined table extraction, regex-based text parsing, duplicate consolidation, invalid-table detection, structured JSON generation, and Markdown/CSV metric reports.
    \item Exposed the pipeline through a REST API and standalone benchmark scripts for batch processing and timing analysis.
\end{itemize}

\resumeProjectHeading
{SAT Practice Platform}{Next.js, React, TypeScript, Prisma, NextAuth, Tailwind}

\begin{itemize}
    \item Developed a role-based SAT preparation platform for students, teachers, admins, and parents with protected portals and validation states.
    \item Implemented assignment flows, timed practice attempts, autosave, score submission, explanations, full practice tests, user invitations, password reset, and parent consent workflows.
    \item Used Prisma with SQLite, NextAuth credentials/JWT, React Hook Form, Zod validation, Radix UI components, TanStack Table, KaTeX, and email notifications.
\end{itemize}

\resumeProjectHeading
{Legal Researcher for Peruvian Telecom Regulation}{Node.js, Express, SQLite, Cron, PDF processing}

\begin{itemize}
    \item Built a daily scraper for El Peruano and OSIPTEL legal norms related to telecommunications, consumer protection, mobile services, SIM cards, and regulatory projects.
    \item Implemented source-specific filters, PDF download, SHA-256 tracking, SQLite persistence, scheduled execution, HTTP API access, email reports, and webhook delivery.
    \item Wrote automated tests using Node's native test runner and kept source keywords centralized for maintainable regulatory matching.
\end{itemize}

\resumeProjectHeading
{Image Recognition REST and Scheduler APIs}{Java, Spring Boot, MySQL, S3, JWT}

\begin{itemize}
    \item Implemented REST APIs for image recognition workflows, verification rules, users, security, image formats, scheduler operations, and external-system integrations.
    \item Used Spring Boot, Spring Security, JPA, validation, OpenAPI documentation, JWT authentication, MySQL, S3 integration, Apache POI, and layered endpoint documentation.
\end{itemize}

\resumeProjectHeading
{Optimization and Model Evaluation Experiments}{Python, Java, ML, Metaheuristics}

\begin{itemize}
    \item Implemented ACO-TW, ABC, PSO, gradient-descent-based reasoning, and custom experiment loops for routing, assignment, scheduling, and parameter-search problems.
    \item Evaluated models and algorithms with custom objective functions, classification/regression metrics, repeated runs, convergence analysis, and hyperparameter sensitivity.
\end{itemize}

% ===================== EDUCATION =====================

\section{Education}

\resumeSubheading
{Pontificia Universidad Catolica del Peru}
{[Start Date] -- [End Date]}
{Bachelor's Degree in Informatics Engineering / Computer Engineering}
{Lima, Peru}

\begin{itemize}
    \item Relevant Coursework: Artificial Intelligence, Software Engineering, Operating Systems, Information Security, Enterprise Architecture, ERP, Business Intelligence, Simulation, Statistics, Data Visualization.
    \item Research Focus: document intelligence, Spanish-language AI systems, RAG evaluation, OCR validation, model efficiency, and applied optimization.
\end{itemize}

% ===================== CERTIFICATIONS =====================

\section{Certifications and Training}

\begin{itemize}
    \item RAG, LLM evaluation, prompt engineering and lightweight AI deployment -- applied project experience
    \item Docker, service orchestration, API gateways, queues and backend observability -- applied project experience
    \item Machine learning, deep learning, OCR, computer vision and optimization -- academic and project experience
    \item Hyperledger Fabric, ISO 20022 and digital asset infrastructure -- independent specialization
\end{itemize}

\end{document}
